---
title: "R Notebook"
output: word_document
---

#Controls Overview
Two types of controls (extraction and sequencing) were used in this experiment, for a total of 10 different controls:
 Extraction Controls (Artificial Colony, Robogut, Zymobiomics Control 1, Zymobiomics Control2, Extraction Blank)
 Sequencing Controls (ATTC 1-4, Zymobiomics Seq Control)
 
#Libraries
```{r}
#Load required libraries
library("biom")
library(data.table)
```

#Control Makeup
Read in each of the controls composition files
```{r}
control_full <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fresh Fecal Optimization_2017.08\\Controls\\Compositions\\MasterSpeciesList.txt",sep="\t",header=TRUE)

#Separate database by control colonies
art.col <- subset(control_full, !(Art.Col=='A'))
zymo.ext200 <- subset(control_full, !(Zymo.Ext=='A'))
zymo.ext2000 <-subset(control_full, !(Zymo.Ext=='A'))
msa1000 <-subset(control_full, !(MSA1000=='A'))
msa1001 <- subset(control_full, !(MSA1001=='A'))
msa1002 <- subset(control_full, !(MSA1002=='A'))
msa1003 <- subset(control_full, !(MSA1003=='A'))
zymo.seq <-subset(control_full, !(zymo.seq=='A'))

#Create list of all the species to review
species_list_controls <- as.character(unique(control_full$Species))
species_list_controls
```

#Sample Data
Read in the entire sample and OTU databses, then merge the databases. Finally filter for controls only
```{r}
#Read in database with variable information and list of samples that did not meet the sequencing threshold
complete_info <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fresh Fecal Optimization_2017.08\\Analysis\\Run1&2\\Manifests\\sampledata.txt",sep="\t",header=TRUE)
complete_removed <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fresh Fecal Optimization_2017.08\\Analysis\\Run1&2\\samples_being_ignored.txt",sep="\t")

#Remove failed samples from info database
filtered_info <-subset(complete_info, !(NephID %in% complete_removed$V1))

#Verify the numbers are the same
nrow(complete_info) - nrow(complete_removed)
nrow(filtered_info)

#Remove Study samples from database
control_info <- subset(filtered_info,!Type=='Study')

#Verify the subsetting
nrow(filtered_info) - sum(filtered_info$Type=='Study')
nrow(control_info)
```

Nephele OTU Table
```{r}
#Read in BIOM database output from Nephele and convert it to a data frame
dat <- read_biom("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fresh Fecal Optimization_2017.08\\Analysis\\Run1&2\\otu_table_mc2_w_tax.biom")
otu_table  <- as.data.frame(as.matrix(biom_data(dat)))

#Transpose the otu_table so same names are in the rows, and species are in the columns
otu_table_flipped <- transpose(otu_table)
colnames(otu_table_flipped) <- rownames(otu_table)
rownames(otu_table_flipped) <- colnames(otu_table)

#Subset the OTU table to include only passed control samples by finding selecting only the rows with names that 
#match the SeqID names of the control_info database
otu_controls <- subset(otu_table_flipped, rownames(otu_table_flipped) %in% control_info$NephID)

#Review database
head(otu_controls)

#Samples in the BIOM database must have representative species with higher than 2 OTU sequences, otherwise they are dropped. In this instance 1 sample was lost - Sample SC326763.PC07578.A12
nrow(otu_controls)
```

Update species names to reference database
```{r}
#Read in reference file
references_complete <- observation_metadata(dat)
head(references_complete)

#Create a list of unique reference species, then remove D_6__
species_list_reference<- unique(references_complete$taxonomy7)
species_list_reference<- gsub("D_6__","",species_list_reference)

#Determine which control species are not included in reference database
missing_species <- species_list_controls[!(species_list_controls %in% species_list_reference)]
missing_species

#Subset the taxonomy in question
temp <- subset(references_complete, taxonomy7=='D_6__Acinetobacter baumannii')
temp
unique(temp$taxonomy7)
```

Generate info and otu table with selected controls
```{r}
#Subset info database with only control information
control_removed 
control_otu 


```