---
title: "Fecal_Genus"
author: "NHI"
date: "July 3, 2018"
output: word_document
editor_options: 
  chunk_output_type: console
---

---
title: "R Notebook"
output: word_document
editor_options: 
  chunk_output_type: console
---

#1. Load required libraries
```{r}
library("biom")
library(data.table)
library(ggplot2)
```

#2. Taxonomy Reference - Control Annotations
Read in controls taxonomy file, grouped by Genus
```{r}
genus_percent <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fresh Fecal Optimization_2017.08\\Controls\\Taxonomy\\Taxonomy_Genus.txt",sep="\t",header=TRUE)

#Add row names to genus full
genus_list <- unique(genus_percent[,"Genus"])
rownames(genus_percent) <- genus_list

#Create Control Composition Database, removing rows of species not in samples, followed by samples not of specified control type
##Two concentrations of zymo sequencing controls
zymo.seq200.con <- subset(genus_percent, !(Zymo.Seq=='NA'))
zymo.seq200.con <- zymo.seq200.con[,c(1:9,17)]
zymo.seq2000.con <- zymo.seq200.con

#ATTC Controls
msa1000.con <-subset(genus_percent, !(MSA1000=='NA'))
msa1000.con <- msa1000.con[,c(1:9,12)]

msa1001.con <- subset(genus_percent, !(MSA1001=='NA'))
msa1001.con <- msa1001.con[,c(1:9,13)]

msa1002.con <- subset(genus_percent, !(MSA1002=='NA'))
msa1002.con <- msa1002.con[,c(1:9,14)]

msa1003.con <- subset(genus_percent, !(MSA1003=='NA'))
msa1003.con <- msa1003.con[,c(1:9,15)]

```

#3. Variable info
Read in the entire control variable database, and controls that failed threshold QC. 
Filter database removing all failed samples, and any that are not a control
```{r}
#Read in database with variable information and list of samples that did not meet the sequencing #threshold
variable_info <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fresh Fecal Optimization_2017.08\\Analysis\\Run1&2\\Manifests\\sampledata.txt",sep="\t",header=TRUE)
samples_failed <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fresh Fecal Optimization_2017.08\\Analysis\\Run1&2\\TaxaCounts\\samplesfailed_nephele1.txt",sep="\t")

#Remove failed samples from info database
filtered_info <-subset(variable_info, !(NephID %in% samples_failed$V1))

#Verify filtering - numbers should match
nrow(variable_info) - nrow(samples_failed)
nrow(filtered_info)

#Remove Study samples from database
control_info <- subset(filtered_info,!Type=='Study')
control_info <- subset(control_info,!Type=='EB')
control_info <- subset(control_info,!Type=='RG')

#Verify the subsetting - numbers should match
nrow(filtered_info) - sum(filtered_info$Type=='Study') - sum(filtered_info$Type=='EB') - sum(filtered_info$Type=='RG')
nrow(control_info)

#Add Row names to dataframe
row.names(control_info) <- control_info$NephID
```

#4. OTU Counts - OTU's by Genus (raw data)
Read in OTU information, and determine total Genus counts
```{r}
#Read in OTU database at Genus level
genus_otu <- read.csv("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fresh Fecal Optimization_2017.08\\Analysis\\Run1&2\\TaxaCounts\\nephele1_taxa_counts_Genus.csv", row.names = 1)

#Create a total row for each sample
sample_list <- colnames(genus_otu)
genus_list <- rownames(genus_otu)

#For each sample
for (a in sample_list){
 #Set sum to 0
 sum=0
 
 #For each genus
 for (b in genus_list){
  #Add the number of counts together
  sum <- genus_otu[b,a] + sum
 }
 
 #When complete, push to a "Total_Reads" row
 genus_otu["Total_Reads",a] <- sum
}
```

#5. OTU Counts - Analysis Genus Expected in Controls by Control Type
Read in OTU information, and determine Genus counts (presence/absence) for each sample
```{r}
#Create control sample list, control type list
sample_list <- unique(control_info$NephID)
control_types <- unique(control_info$Type)

#Create a missing list for non-annotated Genus
missing_anno <- list()

#Determine which genus_list to use
##Run through each sample
for (a in sample_list){
 
 #Set count to 0
 count_present = 0
 count_absent = 0
 
 #Create genus list of control types
 if(control_info[a,"Type"]=="MSA-1000"){
    genus_list <- unique(msa1000.con$Genus)
  } else if(control_info[a,"Type"]=="MSA-1001"){
   genus_list <- unique(msa1001.con$Genus)
   } else if(control_info[a,"Type"]=="MSA-1002"){
    genus_list <- unique(msa1002.con$Genus)
    } else if(control_info[a,"Type"]=="MSA-1003"){
     genus_list <- unique(msa1003.con$Genus)
     } else if(control_info[a,"Type"]=="Zymo.Seq.200"){
      genus_list <- unique(zymo.seq200.con$Genus)
      } else if(control_info[a,"Type"]=="Zymo.Seq.2000"){
       genus_list <- unique(zymo.seq2000.con$Genus)
       } else {genus_list <- data.frame()}
 
 #Convert to character
 genus_list <- as.character(genus_list)
 
 #For each genus is list, determine number of reads
 for (b in genus_list){
  value <- genus_otu[b,a]

  #If value is NA (meaning that the genus was not found in sample), skip
  #If reads are >0 then count_presence increases by 1
  #If reads are =0 then count_absence increases by 1 

  if(is.na(value) | is.null(value)) {
   missing_anno[[b]] <- "No Annotations"
   next
  } else if (value>0) {
   count_present = count_present +1
  } else if (value==0){
   count_absent = count_absent +1
  } else{
   next
   }
 }
 
 #Add final count and length of the list to Control_Info dataframe
 control_info[a,"GenusCount"] <- count_present
 control_info[a,"GenusAbsent"] <- count_absent
 control_info[a,"GenusExpect"] <- length(genus_list)
}

#New Database to store final Genus information
genus_otucount_controls <- data.frame()

#For each of the control types
for (b in control_types){
 
 #Reset counts
 genus=0
 total=0
 count=0
 
 #For each of the samples in the sample list
 for(a in sample_list){
  
  #Create control type groups
  valid <- control_info[a,"Type"]
  
  #If the current samples control type matches the control type of the iteration
  if(valid==b){
   
   #Add the total number of genus found
   #Add the total number of genus expected, and subtract the number missing (this is necessary as some
   #genus are not found in the list, and not necessarily absent from the sample)
   genus <- genus + control_info[a,"GenusCount"]
   total <- total + control_info[a,"GenusCount"] + control_info[a,"GenusAbsent"]
   expect <- control_info[a,"GenusExpect"]
   count <- count + 1
  } else{
    next
   }
 }
 
 #Find the average of the genus and totals
 genus_average <- genus/count
 total_average <- total/count
 
 #Push the totals to the new database 
 genus_otucount_controls[1,b] <- genus_average
 genus_otucount_controls[2,b] <- total_average
 genus_otucount_controls[3,b] <- expect

 #Add Row names
 row.names(genus_otucount_controls) <- c("Genus Count","Total", "Expected")
 genus_otucount_controls
}
```

#6. OTU Counts - Analysis Genus OTU Presence/Absence by Control Type by Plate (Seq Control Only)
Sequencing controls by control type by plate
```{r}
#Unique plates
plates <- unique(control_info$Seq.Plate)

#Control Types for Extraction only - no information for Sequencing controls
control_types <- c("MSA-1000","MSA-1001", "MSA-1002","MSA-1003", "Zymo.Seq.200","Zymo.Seq.2000")

#New Database for genus level count information from seq plates
genus_otucount <- data.frame()

#Reset counters
sums =0
row_count =0
row_count2 =0

#For each of the sequencing plates
for (a in plates){
 
 #Group by control types
 for (b in control_types) {
  
  #Reset counters
  genus =0
  total=0
  expect=0

  #Create a list of samples that are of the specified plate and control type
  sample_list<- subset(control_info, Seq.Plate==a & Type==b)
  sample_list <- unique(sample_list$NephID)

  #For each sample in the list
  for (c in sample_list){
   
   #Count the number of unique genus found
   #Count the total number of gensus matched in reference database
   #Count the number of expected genus for this sample
   genus <- genus + control_info[c,"GenusCount"]
   total <- total + control_info[c,"GenusCount"] + control_info[c,"GenusAbsent"]
   expect <- control_info[c,"GenusExpect"]
  }
  
  #Take the average of all extracted iterations of this control  with this kit of
  genus_average <- genus/length(sample_list)
  total_average <- total/length(sample_list)
  
  #Move the row counters by 1 or 2
  row_count <- sums +1
  row_count2 <- sums+2
  
  #Push information to the database
  genus_otucount[row_count, b] <- genus_average
  genus_otucount[row_count2, b] <- genus_average
 }
 
 #Increase the row counter to skip the two filled rows
 sums = sums+2
}
genus_otucount
#Add row names to the database
row.names (genus_otucount) <- c("PC04924_Count", "PC04924_Expected", "PC04925_Count", "PC04925_Expected", "PC07578_Count", "PC07578_Expected")  
```

#7. OTU Percent - Analysis Genus OTU Abundance (All Samples)
Percentages of all genus, listed by control type
```{r}
#Create sample list
sample_list <- unique(control_info$NephID)
genus_list <- rownames(genus_otu[1:nrow(genus_otu)-1,])

#Create new table
genus_otuabund_sample = data.frame()

#Determine which genus_list to use
##Run through each sample
for (a in sample_list){
 
 for (b in genus_list){
  
  #Determine the reads for the genus, and the total reads for the sample
  reads <- genus_otu[b,a]
  total <- genus_otu["Total_Reads",a]

  #If value is NA (meaning that the genus was not found in sample), skip
  #If reads are >0 then add the read count to reads_total

  if(is.na(reads) | is.null(reads)) {
   next
  } else {
   percent <- (reads/total)
   
   #Add out count percent to the genus_otuabund_sample database
   genus_otuabund_sample[b,a] <- percent
  }
 }
}

#Output the table of all samples
write.csv(genus_otuabund_sample,"Data/genus_otuabund_seq_neph_sample.csv")
```

#8, OTU Percent - Analysis Genus OTU Abundance by Control Type
Percentages of genus included in controls, listed by control type
```{r}
#Unique Types of controls and all genus
control_types <- unique(control_info$Type)
genus_list <- rownames(genus_otu[1:nrow(genus_otu)-1,])

#New Database to store final Genus information
genus_otuabund_control <- data.frame()

#For each of the control types
for (a in control_types){
 
  #For each Genus in list
  for (b in genus_list){
   
   #Reset Counter
   reads <- 0
   total <- 0
   
   #Create a list of samples that are of the specified control type
   sample_list<- subset(control_info, Type==a)
   sample_list <- unique(sample_list$NephID)
   
   #For each of the samples in the sample list
   for(c in sample_list){
    
    #Add up the total percentages from otupercent_sample table [genus,sampleID]
    reads <- genus_otu [b,c] + reads
    total <- genus_otu["Total_Reads",c] + total
   }
   
   #average of genus level by taking total number of reads of that genus divded by 
   #total number of reads for samples included
   #Push value to new dataframe
   percent <- reads / total
   genus_otuabund_control[b,a] <- percent
  }
 }

#Output to text file
write.csv(genus_otuabund_control,"Data/genus_otuabund_seq_neph_control.csv")
```

#9. OTU Percent - Analysis Genus OTU Abundance by Control Type by Plate (Seq only)
```{r}
#All known genus (removes uncultured group and total_reads)
genus_list <- rownames(genus_otu[1:(nrow(genus_otu)-2),])
#Store column number for final dataframe
cols <- ncol(genus_percent)

#For each of the seq plates
for (a in plates){
 
  #For each Genus in list
  for (b in genus_list){
   
   #Reset Counter
   reads <- 0
   total <- 0
   
   #Create a list of samples that are of the specified control type
   sample_list<- subset(control_info, Seq.Plate==a)
   sample_list <- unique(sample_list$NephID)
   
   #For each of the samples in the sample list
   for(c in sample_list){
    
    #Add up the total percentages from otupercent_sample table [genus,sampleID]
    reads <- genus_otu [b,c] + reads
    total <- genus_otu["Total_Reads",c] + total
   }
   
   #average of genus level by taking total number of reads of that genus divded by 
   #total number of reads for samples included
   #Push value to new dataframe
   percent <- reads / total
   genus_percent[b,a] <- percent
  }
}
#Repeat at the control by plate level
control_types<-c("MSA-1000", "MSA-1001", "MSA-1002", "MSA-1003", "Zymo.Seq.200", "Zymo.Seq.2000")
hide <- genus_percent

i = ncol(genus_percent) +1
#For each of the kits types
for (a in plates){
 for(b in control_types){
  #Create a list of samples that are of the specified control type
  sample_list<- subset(control_info, Seq.Plate==a)
  sample_list<- subset(sample_list,Type==b)
  sample_list <- unique(sample_list$NephID)
  
  for(c in genus_list){
   
   #Reset Counter
   reads <- 0
   total <- 0

    #For each of the samples in the sample list
   for(d in sample_list){
    
    #Add up the total percentages from otupercent_sample table [genus,sampleID]
    reads <- genus_otu [c,d] + reads
    total <- genus_otu["Total_Reads",d] + total
   }
   
   #average of genus level by taking total number of reads of that genus divded by 
   #total number of reads for samples included
   #Push value to new dataframe
   percent <- reads / total
   genus_percent[c,i] <- percent
  }
  i=i+1
 }
}

i=cols+1
#Add column names
new_colnames<- c("PC04924_OB", "PC04925_OB","PC07578_OB",
                 "PC04924_MSA1000", "PC04924_MSA1001", "PC04924_MSA1002", "PC04924_MSA1003", 
                 "PC04924_Seq200", "PC04924_Seq2000",
                 "PC04925_MSA1000", "PC04925_MSA1001","PC04925_MSA1002","PC04925_MSA1003", 
                 "PC04925_Seq200", "PC04925_Seq2000",
                 "PC07578_MSA1000", "PC07578_MSA1001","PC07578_MSA1002","PC07578_MSA1003", 
                 "PC07578_Seq200", "PC07578_Seq2000")
#for each header name, apply it to the recently created columns
for (a in new_colnames){
 colnames(genus_percent)[i]<- a
 i=i+1
}

#Output to text file
write.csv(genus_percent,"Data/genus_outabund_seq_neph.csv")
```

#10.OTU Percent - Analysis of Genus Percent Abundance Differences
Expected versus observed by plate (Seq Only) - Create new genus databse with only MSA and Zymo Seq Columns
```{r}
#Remove genus groups not included in control types
hide<-genus_percent
temp <- data.frame()
colnames(genus_percent)
temp <- genus_percent[,c(12:15,17,21:38)]
temp <- subset(temp, !('MSA-1000'=="NA" & 'MSA-1001'=="NA"& 'MSA-1002'=="NA"& 'MSA-1003'=="NA"& 
                       'Zymo.Seq.200'=="NA"& 'Zymo.Seq.2000'=="NA"))
temp2 <- temp[,c(1:5,5)]
head(temp2)

#Save the column start
cols=ncol(genus_percent)
cols

#Set the counter
i=cols+1
counter=0

#Create lists to compare plates
genus_list <- unique(rownames(temp))
compare_list <- colnames(temp[,6:23])

#For each control 
for (a in compare_list){
 if(counter==6){
  counter==1
 } else{counter=counter+1}
 
 #For each genus, within the specific colony list
 for (b in genus_list){
  
  #Determine abundance value
  value <- temp2[b,counter]
  
  #If the value is NA fill in NA
  if(is.na(value)){
   genus_percent[b,i] <- "NA"
  }
  #Otherwise, fill in the abundance rate
  else{
   genus_percent[b,i] <- ((value - temp[b,a])/value)*100*-1
  }
  
 }
 #increase the counter
 i=i+1
}

#Add column names
new_colnames <- c("PC04924-MSA1000_Diff", "PC04924-MSA1001_Diff", "PC04924-MSA1002_Diff",
                  "PC04924-MSA1003_Diff", "PC04924-Seq.200_Diff", "PC04924-Seq.2000_Diff",
                  "PC04925-MSA1000_Diff", "PC04925-MSA1001_Diff", "PC04925-MSA1002_Diff",
                  "PC04925-MSA1003_Diff", "PC04925-Seq.200_Diff", "PC04925-Seq.2000_Diff",
                  "PC07578-MSA1000_Diff", "PC07578-MSA1001_Diff", "PC07578-MSA1002_Diff",
                  "PC07578-MSA1003_Diff", "PC07578-Seq.200_Diff", "PC07578-Seq.2000_Diff")

i=cols+1
for (a in new_colnames){
 colnames(genus_percent)[i]<- a
 i=i+1
}

#subset table to include only relevant genus, and only seq control information
colnames(genus_percent)
temp <- genus_percent[,c(12:15,17,21:56)]
temp <- subset(temp, !('MSA-1000'=="NA" & 'MSA-1001'=="NA"& 'MSA-1002'=="NA"& 'MSA-1003'=="NA"& 
                       'Zymo.Seq.200'=="NA"& 'Zymo.Seq.2000'=="NA"))

#Output file
write.csv(temp,"Data/genus_otuabund_seq_neph_obsVSexp.csv")

#Create tableau file
temp2<-data.frame()
count=1
for (a in 24:ncol(temp)){
 for(b in 1:nrow(temp)){
  temp2[count,"Genus"] <- rownames(temp[b,])
  temp2[count,"value"] <- temp[b,a]
  temp2[count,"Type"] <- colnames(temp[a])
  count=count+1
 }
}
head(temp2)
write.csv(temp2,"Data/genus_otuabund_seq_neph_obsVexp_tab.csv")
```

#11.OTU Percent - Not expected Genus by Control Type by Plate (Seq Only)
```{r}
#Create temp dataframe from genus_percent
colnames(genus_percent)
temp <- genus_percent[,c(12:15,17,21:56)]

#Subset by those Genus not expected to be in extraction
temp <- subset(temp, is.na(MSA1000) | is.na(MSA1001) | is.na(MSA1002) | 
                is.na(MSA1003) | is.na(Zymo.Seq))
#Output the unexpected percent frequency database
write.csv(temp,"Data/genus_otuabund_seq_neph_other.csv")
colnames(temp)
head(temp)

#Create tableau file
temp2<-data.frame()
count=1
for (a in 6:23){
 for(b in 1:nrow(temp)){
  temp2[count,"Genus"] <- rownames(temp[b,])
  temp2[count,"value"] <- temp[b,a]
  temp2[count,"Type"] <- colnames(temp[a])
  count=count+1
 }
}
head(temp2)
write.csv(temp2,"Data/genus_otuabund_seq_neph_other_tab.csv")
```

#12 Plot graph - Actual to Expected
```{r}
###############
#MSA1000
###############
#Create dataframe
colnames(genus_percent)
msa1000.plates <- genus_percent[,c("MSA1000","Genus","PC04924_MSA1000","PC04925_MSA1000","PC07578_MSA1000")]
msa1000.plates <- subset(msa1000.plates,!(is.na(MSA1000)))
rownames(msa1000.plates)
colnames(msa1000.plates)

#Change formatting for graphics
temp3<-data.frame()
count=1
for (a in c(3:5)){
 count2=1
 for(b in 1:nrow(msa1000.plates)){
  temp3[count,"Expected"]<- msa1000.plates[count2,"MSA1000"]
  temp3[count,"Genus"] <- rownames(msa1000.plates[b,])
  temp3[count,"value"] <- msa1000.plates[b,a]
  temp3[count,"Type"] <- colnames(msa1000.plates[a])
  count=count+1
  count2=count2+1
 }
}
head(temp3)

#Plot graph
ggplot(data=temp3,mapping=aes(x=Expected,y=value,color=Type))+
 #geom_point()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="MSA1000-Nephele", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(x=Expected,y=value,label=Genus),check_overlap = TRUE)


###############
#MSA1001
###############
#Create dataframe
colnames(genus_percent)
msa1001.plates <- genus_percent[,c("MSA1001","Genus","PC04924_MSA1001","PC04925_MSA1001","PC07578_MSA1001")]
msa1001.plates <- subset(msa1001.plates,!(is.na(MSA1001)))
rownames(msa1001.plates)
colnames(msa1001.plates)

#Change formatting for graphics
temp3<-data.frame()
count=1
for (a in c(3:5)){
 count2=1
 for(b in 1:nrow(msa1001.plates)){
  temp3[count,"Expected"]<- msa1001.plates[count2,"MSA1001"]
  temp3[count,"Genus"] <- rownames(msa1001.plates[b,])
  temp3[count,"value"] <- msa1001.plates[b,a]
  temp3[count,"Type"] <- colnames(msa1001.plates[a])
  count=count+1
  count2=count2+1
 }
}
temp3

#Create databases of upper and lower expected frequencies
temp4 <- subset(temp3,Expected>.2)
temp4

temp5 <- subset(temp3,Expected<.2)
temp5

#Plot graphs
##All data
ggplot(data=temp3,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_line()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="MSA1001-Nephele", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))
##Uppers
ggplot(data=temp4,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_point()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="MSA1001-Nephele-Upper", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))
##Lowers
ggplot(data=temp5,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_point()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="MSA1001-Nephele-Lower", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))

###############
#MSA1002
###############
#Create dataframe
colnames(genus_percent)
msa1002.plates <- genus_percent[,c("MSA1002","Genus","PC04924_MSA1002","PC04925_MSA1002","PC07578_MSA1002")]
msa1002.plates <- subset(msa1002.plates,!(is.na(MSA1002)))
rownames(msa1002.plates)
colnames(msa1002.plates)

#Change formatting for graphics
temp3<-data.frame()
count=1
for (a in c(3:5)){
 count2=1
 for(b in 1:nrow(msa1002.plates)){
  temp3[count,"Expected"]<- msa1002.plates[count2,"MSA1002"]
  temp3[count,"Genus"] <- rownames(msa1002.plates[b,])
  temp3[count,"value"] <- msa1002.plates[b,a]
  temp3[count,"Type"] <- colnames(msa1002.plates[a])
  count=count+1
  count2=count2+1
 }
}
temp3

#Plot graphs
##All data
ggplot(data=temp3,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_line()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="MSA1002-Nephele", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))

###############
#MSA1003
###############
#Create dataframe
colnames(genus_percent)
msa1003.plates <- genus_percent[,c("MSA1003","Genus","PC04924_MSA1003","PC04925_MSA1003","PC07578_MSA1003")]
msa1003.plates <- subset(msa1003.plates,!(is.na(MSA1003)))
rownames(msa1003.plates)
colnames(msa1003.plates)

#Change formatting for graphics
temp3<-data.frame()
count=1
for (a in c(3:5)){
 count2=1
 for(b in 1:nrow(msa1003.plates)){
  temp3[count,"Expected"]<- msa1003.plates[count2,"MSA1003"]
  temp3[count,"Genus"] <- rownames(msa1003.plates[b,])
  temp3[count,"value"] <- msa1003.plates[b,a]
  temp3[count,"Type"] <- colnames(msa1003.plates[a])
  count=count+1
  count2=count2+1
 }
}
unique(temp3$Expected)
temp3

#Create databases of upper and lower expected frequencies
temp4 <- subset(temp3,Expected>.1)
temp4

temp5 <- subset(temp3,Expected<.2)
temp5

#Plot graphs
##All data
ggplot(data=temp3,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_line()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="MSA1003-Nephele", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))
##Uppers
ggplot(data=temp4,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_point()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="MSA1003-Nephele-Upper", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))
##Lowers
ggplot(data=temp5,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_point()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="MSA1003-Nephele-Lower", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))

###############
#SEQ200
###############
#Create dataframe
colnames(genus_percent)
seq200.plates <- genus_percent[,c("Zymo.Seq","Genus","PC04924_Seq200","PC04925_Seq200","PC07578_Seq200")]
seq200.plates <- subset(seq200.plates,!(is.na(Zymo.Seq)))
rownames(seq200.plates)
colnames(seq200.plates)

#Change formatting for graphics
temp3<-data.frame()
count=1
for (a in c(3:5)){
 count2=1
 for(b in 1:nrow(seq200.plates)){
  temp3[count,"Expected"]<- seq200.plates[count2,"Zymo.Seq"]
  temp3[count,"Genus"] <- rownames(seq200.plates[b,])
  temp3[count,"value"] <- seq200.plates[b,a]
  temp3[count,"Type"] <- colnames(seq200.plates[a])
  count=count+1
  count2=count2+1
 }
}
temp3
unique(temp3$Expected)

#Plot graphs
##All data
ggplot(data=temp3,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_line()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="Zymo Seq 200-Nephele", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))
###############
#SEQ2000
###############
#Create dataframe
colnames(genus_percent)
seq2000.plates <- genus_percent[,c("Zymo.Seq","Genus","PC04924_Seq2000","PC04925_Seq2000","PC07578_Seq2000")]
seq2000.plates <- subset(seq2000.plates,!(is.na(Zymo.Seq)))
rownames(seq2000.plates)
colnames(seq2000.plates)

#Change formatting for graphics
temp3<-data.frame()
count=1
for (a in c(3:5)){
 count2=1
 for(b in 1:nrow(seq2000.plates)){
  temp3[count,"Expected"]<- seq2000.plates[count2,"Zymo.Seq"]
  temp3[count,"Genus"] <- rownames(seq2000.plates[b,])
  temp3[count,"value"] <- seq2000.plates[b,a]
  temp3[count,"Type"] <- colnames(seq2000.plates[a])
  count=count+1
  count2=count2+1
 }
}
temp3
unique(temp3$Expected)

#Plot graphs
##All data
ggplot(data=temp3,mapping=aes(x=Expected,y=value,color=Type,jitter(Expected,factor=1)))+
 #geom_line()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="Zymo Seq 2000-Nephele", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(label=Genus))
```