---
title: "Fecal Extractions"
output: word_document
editor_options: 
  chunk_output_type: console
---
Based on vigenette
https://benjjneb.github.io/dada2/bigdata_paired.html
https://bioconductor.org/packages/devel/bioc/vignettes/dada2/inst/doc/dada2-intro.html

Install packages
```{r}
.cran_packages <- c("ggplot2", "gridExtra", "survival")
.bioc_packages <- c("dada2", "phyloseq", "DECIPHER", "phangorn")
.inst <- .cran_packages %in% installed.packages()
if(any(!.inst)) {
   install.packages(.cran_packages[!.inst])
}
.inst <- .bioc_packages %in% installed.packages()
if(any(!.inst)) {
   source("http://bioconductor.org/biocLite.R")
   biocLite(.bioc_packages[!.inst], ask = F)
}
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE)
source("http://bioconductor.org/biocLite.R")
biocLite("phyloseq")
```


Load Library's
```{r}
library("knitr")
library(dada2); packageVersion("dada2")
library(phangorn)
library("phyloseq")
library("gridExtra")
library("ggplot2")
```

Perform Filtering of data of Run 3
```{r}
# Set path of samples
pathF <- "T:/DCEG/CGF/Laboratory/Projects/MR-0084/NP0084-MB5/QC Data/qiime/Combo with NP0084-MB4/04_23_18_input_combo/Run3/forward"
pathR <- "T:/DCEG/CGF/Laboratory/Projects/MR-0084/NP0084-MB5/QC Data/qiime/Combo with NP0084-MB4/04_23_18_input_combo/Run3/reverse"

#Set path of filtered samples
filtpathF <- file.path(pathF, "filtered") # Filtered forward files go into the pathF/filtered/ subdirectory
filtpathR <- file.path(pathR, "filtered") # ...
fastqFs <- sort(list.files(pathF, pattern="fastq.gz"))
fastqRs <- sort(list.files(pathR, pattern="fastq.gz"))

#QC check to make sure fastq forward and reverse are the same
if(length(fastqFs) != length(fastqRs)) stop("Forward and reverse files do not match.")

#Review Quality Profiles of any file
plotQualityProfile(filtpathF[1])
plotQualityProfile(filtpathR)


# Filt and trim the data, store new files in filtered path- forward trimmed LESS than reverse
filterAndTrim(fwd=file.path(pathF, fastqFs), filt=file.path(filtpathF, fastqFs),
              rev=file.path(pathR, fastqRs), filt.rev=file.path(filtpathR, fastqRs),
              truncLen=c(240,200), maxEE=2, truncQ=11, maxN=0, rm.phix=TRUE,
              compress=TRUE, verbose=TRUE, multithread=TRUE)
```

Construct Matrix for Run 3
```{r}
# File parsing of filtered FastQ files
filtpathF <- "T:/DCEG/CGF/Laboratory/Projects/MR-0084/NP0084-MB5/QC Data/qiime/Combo with NP0084-MB4/04_23_18_input_combo/Run3/forward/filtered" 
filtpathR <- "T:/DCEG/CGF/Laboratory/Projects/MR-0084/NP0084-MB5/QC Data/qiime/Combo with NP0084-MB4/04_23_18_input_combo/Run3/reverse/filtered"
filtFs <- list.files(filtpathF, pattern="fastq.gz", full.names = TRUE)
filtRs <- list.files(filtpathR, pattern="fastq.gz", full.names = TRUE)

#Parse sample names
sample.namesF <- sapply(strsplit(basename(filtFs), "_"), `[`, 1) # Assumes filename = samplename_primer.fastq.gz
sample.namesR <- sapply(strsplit(basename(filtRs), "_"), `[`, 1) # Assumes filename = samplename_primer.fastq.gz
if(!identical(sample.namesF, sample.namesR)) stop("Forward and reverse files do not match.")

#Remove unneeded files
#Apply names - are the same, so apply F to both - to the filtered files
names(filtFs) <- sample.namesF
names(filtRs) <- sample.namesF

# Learn error rates
set.seed(100)
errF <- learnErrors(filtFs, nbases=1e8, multithread=TRUE)
errR <- learnErrors(filtRs, nbases=1e8, multithread=TRUE)

# Sample inference and merger of paired-end reads
mergers <- vector("list", length(sample.namesF))
names(mergers) <- sample.namesF
for(sam in sample.namesF) {
  cat("Processing:", sam, "\n")
    derepF <- derepFastq(filtFs[[sam]])
    ddF <- dada(derepF, err=errF, multithread=TRUE)
    derepR <- derepFastq(filtRs[[sam]])
    ddR <- dada(derepR, err=errR, multithread=TRUE)
    merger <- mergePairs(ddF, derepF, ddR, derepR)
    mergers[[sam]] <- merger
}

#Stopped at SC304924-PC07578-A-11

rm(derepF); rm(derepR)

# Construct sequence table and write
seqtab <- makeSequenceTable(mergers)
saveRDS(seqtab, "T:/DCEG/CGF/Laboratory/Projects/MR-0084/NP0084-MB5/QC Data/qiime/Combo with NP0084-MB4/04_23_18_input_combo/Run3/Output/seqtabrun3.rds")
```

Chimeras and Taxonomy - Merge both runs
```{r}
# Merge multiple runs 
st1 <- readRDS("C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/Output/seqtab073.rds")
st2 <- readRDS("C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/Output/seqtab074.rds")
st.all <- mergeSequenceTables(st1, st2)

# Remove chimeras
seqtab <- removeBimeraDenovo(st.all, method="consensus", multithread=TRUE)

# Assign taxonomy
taxtab <- assignTaxonomy(seqtab, "C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/References", multithread=TRUE)

# Write to disk
saveRDS(seqtab, "C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/Output/seqtab_final.rds") 
saveRDS(tax, "C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/Output/tax_final.rds")
```

Construct Phylo Tree
```{r}
#Pull all the sequence data
seqs <- getSequences(seqtab)
names(seqs) <- seqs 
alignment <- AlignSeqs(DNAStringSet(seqs), anchor=NA)

#Allign the tree
phang.align <- phyDat(as(alignment, "matrix"), type="DNA")
dm <- dist.ml(phang.align)
treeNJ <- NJ(dm) 
fit = pml(treeNJ, data=phang.align)

fitGTR <- update(fit, k=4, inv=0.2)
fitGTR <- optim.pml(fitGTR, model="GTR", optInv=TRUE, optGamma=TRUE,
                      rearrangement = "stochastic", control = pml.control(trace = 0))
detach("package:phangorn", unload=TRUE)
```

Create PhyloSeq Object
```{r}
#Set path of metadata of sample information
mimarks_path <- file.path("C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Input_PhyloSeq.csv")
samdf <- read.csv(mimarks_path, header=TRUE)
all(rownames(seqtab) %in% samdf$SampleID) 
# Verificattion - TRUE - if false check sample names - they must match

#Create phylo object for conversion - includes tax data and metadata
rownames(samdf) <- samdf$SampleID
samdf <- samdf[rownames(seqtab), ]
ps <- phyloseq(tax_table(taxtab),
                 sample_data(samdf),
                 otu_table(seqtab, taxa_are_rows = FALSE),
                 phy_tree(fitGTR$tree))
saveRDS(ps, "C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/Output/ps_ori.rds")
```

Load Phylo Data
```{r}
#load the phylo object
ps = readRDS("C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/Output/ps_ori.rds")
```

PhyloSeq Filtering - Supervised
```{r}
#Determine if anything needs to be filtered
table(tax_table(ps)[, "Phylum"], exclude = NULL) 

#Remove all NA from set
ps0 <- subset_taxa(ps, !is.na(Phylum) & !Phylum %in% c("", "uncharacterized")) 

# Compute prevalence of each feature, store as data.frame
prevdf = apply(X = otu_table(ps0),
                 MARGIN = ifelse(taxa_are_rows(ps0), yes = 1, no = 2),
                 FUN = function(x){sum(x > 0)})
# Add taxonomy and total read counts to this data.frame
prevdf = data.frame(Prevalence = prevdf,
                      TotalAbundance = taxa_sums(ps0),
                      tax_table(ps0))
plyr::ddply(prevdf, "Phylum", function(df1){cbind(mean(df1$Prevalence),sum(df1$Prevalence))})

#Filter species foud in one one sample - Parcubacteria, Saccharibacteria
filterPhyla = c("Parcubacteria", "Saccharibacteria")

# Filter entries with unidentified Phylum.
ps1 = subset_taxa(ps0, !Phylum %in% filterPhyla)

#Review diversity (richness)
plot_richness(ps1,x="Timepoint",measures=c("Shannon","Simpson", "Fisher"), color="Timepoint") + theme_bw() + geom_boxplot()

plot_richness(ps1,x="Diet",measures=c("Shannon","Simpson", "Fisher"), color="Diet") + theme_bw() + geom_boxplot()

plot_richness(ps1,x="SampleType",measures=c("Shannon","Simpson", "Fisher"), color="Diet") + theme_bw() + geom_boxplot()


```

PhyloSeq Filtering - Unsupervised
```{r}
#Review dataset for phylum that are under represented
prevdf1 = subset(prevdf, Phylum %in% get_taxa_unique(ps1, "Phylum"))
ggplot(prevdf1, aes(TotalAbundance, Prevalence / nsamples(ps0),color=Phylum)) + geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) + geom_point(size = 2, alpha = 0.7) + scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") + facet_wrap(~Phylum) + theme(legend.position="none")
```

Define New thresholds
```{r}
#  Define prevalence threshold as 5% of total samples
prevalenceThreshold = 0.05 * nsamples(ps0)
prevalenceThreshold

# Execute prevalence filter, using `prune_taxa()` function
keepTaxa = rownames(prevdf1)[(prevdf1$Prevalence >= prevalenceThreshold)]
ps2 = prune_taxa(keepTaxa, ps0)
```

Aggregate the data
```{r}
# How many genera would be present after filtering?
length(get_taxa_unique(ps2, taxonomic.rank = "Genus"))

#Rank on Genus for set - compare
ps3 = tax_glom(ps2, "Genus", NArm = TRUE)
h1 = 0.4
ps4 = tip_glom(ps2, h = h1)
multiPlotTitleTextSize = 8
p2tree = plot_tree(ps2, method = "treeonly",
                     ladderize = "left",
                     title = "Before Agglomeration") +
  theme(plot.title = element_text(size = multiPlotTitleTextSize))
p3tree = plot_tree(ps3, method = "treeonly",
                     ladderize = "left", title = "By Genus") +
  theme(plot.title = element_text(size = multiPlotTitleTextSize))
p4tree = plot_tree(ps4, method = "treeonly",
                     ladderize = "left", title = "By Height") +
  theme(plot.title = element_text(size = multiPlotTitleTextSize))

# group plots together
plot(grid.arrange(nrow = 1, p2tree, p3tree, p4tree))
```

Transform data for abundances
```{r}
plot_abundance = function(physeq,title = "",
			     Facet = "Order", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(physeq, Phylum %in% c("Firmicutes"))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "Diet",y = "Abundance",
                                 color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
                position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

#Transform and plot the before and after
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})
plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"Genus")

# Combine each plot into one graphic.
grid.arrange(nrow = 2, plotBefore, plotAfter)

#Subset further based on taxa findings, if necessary
psOrd = subset_taxa(ps3, Order == "Lactobacillales")
plot_abundance(psOrd, Facet = "Genus", Color = NULL)
```
Create Phylogeny Trees
```{r}
#Remove decimals from tree for labels
phy_tree(ps3ra)$node.label = substr(phy_tree(ps3ra)$node.label, 1, 4)

#Plot by Variables
plot_tree(ps3ra, nodelabf=nodeplotboot(), ladderize="left", color="Timepoint")
plot_tree(ps3ra, nodelabf=nodeplotboot(), ladderize="left", color="Diet")
plot_tree(ps3ra, nodelabf=nodeplotboot(), ladderize="left", color="SampleType")
plot_tree(ps3ra, nodelabf=nodeplotboot(), ladderize="left", color="Timepoint", shape="Diet")


#Plot by Phylum
plot_tree(ps3ra,color="Phylum", shape="Timepoint")
plot_tree(ps3ra,color="Phylum", shape="Diet")
plot_tree(ps3ra,color="Phylum", shape="SampleType")

#Plot by Class
plot_tree(ps3ra,color="Class", shape="Timepoint")
plot_tree(ps3ra,color="Class", shape="Diet")
plot_tree(ps3ra,color="Class", shape="SampleType")

#Plot by Genus
plot_tree(ps3ra,color="Genus", shape="Timepoint")
plot_tree(ps3ra,color="Genus", shape="Diet")
plot_tree(ps3ra,color="Genus", shape="SampleType")

#Create spirl plots
plot_tree(ps3ra, color="Diet", ladderize="left") + coord_polar(theta="y")

```

Create Barplots by Phylum and Class
```{r}
ps4 <- prune_taxa(taxa_sums(ps3ra) > 0, ps3ra)

plot_bar(ps4,fill="Phylum",x="Worm")
plot_bar(ps4,fill="Phylum",x="Timepoint")
plot_bar(ps4,fill="Phylum",x="SampleType")
plot_bar(ps4,fill="Phylum",x="Diet")
plot_bar(ps4,fill="Class",x="Diet")
plot_bar(ps4,fill="Class",x="SampleType")
plot_bar(ps4,fill="Class",x="Timepoint")

```

Save ps
```{r}
ps_pruned <-saveRDS(ps3ra, "C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/Output/ps_pruned.rds")
ps_filt <- saveRDS(ps2, "C:/Users/sevillas2/Google Drive/My Documents/Education/George Mason University/BINF703/2018_Spring_BaranovaGillevet/Analysis/DADA2/Output/ps_filt.rds")
```

