---
title: "R Notebook"
output: word_document
editor_options: 
  chunk_output_type: console
---
#1. Load required libraries
```{r}
library("biom")
library(data.table)
library(ggplot2)
```

#2. Filtering of samples, control variable information
```{r}
#Read in database with variable information
samples_variables_complete <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Analysis\\ExpA_Run1&2\\Manifests\\sampledata_abrf.txt",sep="\t",header=TRUE)

#Read in database of failed samples
samples_failed <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Analysis\\ExpA_Run1&2\\TaxaCounts\\samplesfailed_q2.txt",sep="\t")

#Remove failed samples from info database
samples_filtered <-subset(samples_variables_complete, !(NephID %in% samples_failed$V1))

#Verify filtering - numbers should match
nrow(samples_variables_complete) - nrow(samples_failed)
nrow(samples_filtered)

#Only include Ext Controls
control_types <- c("RG","Study")
samples_variables_f <- samples_filtered[samples_filtered$Type %in% control_types,]

#Add sample ID's as row names
row.names(samples_variables_f) <- samples_variables_f$NephID
```

#3. OTU Counts by sample
Read in OTU information, and determine total counts for all samples
```{r}
#Read in OTU database at Genus level
samples_otu_complete <- read.csv("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Analysis\\ExpA_Run1&2\\TaxaCounts\\qiime2_taxa_counts_genus_silva_abrf.csv", row.names = 1)

#Create a list of all samples and all genus present in at least one sample
sample_list_complete <- colnames(samples_otu_complete)
genus_found <- rownames(samples_otu_complete)

#For each sample, determine the total read counts
for (a in sample_list_complete){
 #Set sum to 0
 sum=0
 
 #For each genus
 for (b in genus_found){
  #Add the number of counts together
  sum <- samples_otu_complete[b,a] + sum
 }
 
 #When complete, push the sum to a new "Total_Reads" row
 samples_otu_complete["Total_Reads",a] <- sum
}
write.csv(samples_otu_complete,"temp.csv")
```

#4. OTU Counts by Control Type
Determine genus count for each sample, average across control type
```{r}
#Create control sample list, control type list
sample_list_f <- unique(samples_variables_f$NephID)
control_types

#Genus count for each sample 
for (a in sample_list_f){
  count_present=0
  
  for (b in genus_found){
   value <- samples_otu_complete[b,a]
   
   if(value==0) {
    next
   } else{
     count_present = count_present +1
    }
  }
  samples_variables_f[a,"GenusObs"] <- count_present

 }
 
###Testing
write.csv(samples_variables_f,"temp1.csv")


#New Database to store final Genus information
samples_obs_count_type <- data.frame()
kit_types<-unique(samples_variables_f$Kit)

#Start counter
count_present=0

#For each control type
for (a in control_types){
 
 #For each kit
 for (b in kit_types){
  #Create temp database for kit and control type
  temp <- subset(samples_variables_f,Type==a)
  temp <- subset(temp,Kit==b)
  
  #Create sample list
  sample_list<- unique(temp$NephID)
  
  #For each of these samples
  for (c in sample_list){
   
   #Find observed value and add
   value <- temp[c,"GenusObs"]
   count_present=count_present+value
  }
  
  #Create new column and pass average
  col_name <- paste(a,"_",b,sep="")
  samples_obs_count_type["GenObs_Av",col_name]<-count_present/length(sample_list)
  
  #Reset counter
  count_present=0
 }
}
write.csv(samples_obs_count_type,"study_avgenus.csv")
```

*************STOP

#6. OTU Counts Count of Presence/Absence by Control Type by Kit
```{r}
#Create list of kits, removing "None" as a kit type
kit_types <- unique(samples_variables_f$Kit)
kit_types <- kit_types[!kit_types %in% "None"]

#New Database for genus level count information from kits
samples_obs_count_kit <- data.frame()

#Reset counters
sums =0
row_count =0

#For each of the extraction kits
for (a in kit_types){
 
 #For each control type
 for (b in control_types) {
  
  #Reset counters
  count_present =0
  expect=0

  #Create a list of samples that are of the specified kit and control type
  sample_list_temp<- subset(samples_variables_f, Kit==a & Type==b)
  sample_list_temp <- unique(sample_list_temp$NephID)

  #For each sample in the list
  for (c in sample_list_temp){
   
   #Count the number of genus observed (value is those observed that were expected)
   count_present <- count_present + samples_variables_f[c,"GenusObs"]
  }
  
  #Average of all counts of this control  of this kit type
  genus_average <- count_present/length(sample_list_temp)

  #Move the row counters by 1
  row_count <- sums +1

  #Push information to the database
  samples_obs_count_kit[row_count, b] <- genus_average
 }
 
 #Increase the row counter to skip the two filled rows
 sums = sums+1
}

#Add row names to the database
row.names (samples_obs_count_kit) <- c("Qiagen DSPs_Obs", "Zymo Mag Bead_Obs", "PowerMag Soil_Obs", "PowerMag MB_Obs", "Qiagen QIAmp_Obs") 

###Testing
write.csv(samples_obs_count_kit,"temp3.csv")
```

#7. OTU Percent - Analysis Genus OTU Abundance
```{r}
#Create new table
samples_obs_abund = data.frame()

#Determine which genus_expect to use
##Run through each sample
for (a in sample_list_f){
 
 #Check for each of the expected genus
 for (b in genus_expect){
  
  #Determine the reads for the genus, and the total reads for the sample
  reads <- samples_otu_complete[b,a]
  total <- samples_otu_complete["Total_Reads",a]

  #If value is NA (meaning that the genus was not found in sample), skip
  #If reads are >0 then add the read count to reads_total

  if(is.na(reads) | is.null(reads)) {
   next
  } else {
   percent_exp <- (reads/total)
   
   #Add percent to the samples_obs_abund database
   samples_obs_abund[b,a] <- percent_exp
  }
 }
}

#Output the table of all samples
write.csv(samples_obs_abund,"Data/ext_obs_abund.csv")
```

#8. OTU Percent - Analysis Genus OTU Abundance by Control Type
```{r}
#New Database to store final Genus information
samples_obs_abund_vari <- data.frame()

#For each of the control types
for (a in control_types){
 
  #For each Genus in list
  for (b in genus_expect){
   
   #Reset Counter
   reads <- 0
   total <- 0
   
   #Create a list of samples that are of the specified control type
   sample_list_temp<- subset(samples_variables_f, Type==a)
   sample_list_temp <- unique(sample_list_temp$NephID)
   
   #For each of the samples in the sample list
   for(c in sample_list_temp){
    
    #Add up the total percentages from samples_obs_abund [genus,sampleID]
    reads <- samples_otu_complete [b,c] + reads
    total <- samples_otu_complete["Total_Reads",c] + total
   }
   
   #average of genus level by taking total number of reads of that genus divded by total number of reads for samples included
   #Push value to new dataframe
   percent_obs <- reads / total
   samples_obs_abund_vari[b,a] <- percent_obs
  }
 }

#Output to text file
write.csv(samples_obs_abund_vari,"Data/ext_obs_abund_vari.csv")
```

#9. OTU Percent - Analysis Exp OTU Rel Abun by Control by Kit
```{r}
#For each of the kits types
for (a in kit_types){
 
  #For each Genus in list
  for (b in genus_expect){
   
   #Reset Counter
   reads <- 0
   total <- 0
   
   #Create a list of samples that are of the specified control type
   sample_list_temp<- subset(samples_variables_f, Kit==a)
   sample_list_temp <- unique(sample_list_temp$NephID)
   
   #For each of the samples in the sample list
   for(c in sample_list_temp){
    
    #Add total reads from samples_otu_complete [genus,sampleID]
    reads <- samples_otu_complete [b,c] + reads
    total <- samples_otu_complete["Total_Reads",c] + total
   }
   
   #Calc Rel abund (total number of reads / total number of reads)
   #Push value to new dataframe
   percent_obs <- reads / total
   samples_obs_abund_vari[b,a] <- percent_obs
  }
}
#Output to text file
write.csv(samples_obs_abund_vari,"Data/ext_obs_abund_vari.csv")

#For each of the kits types
for (a in kit_types){
 
 #For each of the control types
 for(b in control_types){
  #Create a list of samples that are of the specified control type
  sample_list_temp<- subset(samples_variables_f, Kit==a & Type==b)
  sample_list_temp <- unique(sample_list_temp$NephID)
  
  #For each of the expected genus
  for(c in genus_expect){
   
   #Reset Counter
   reads <- 0
   total <- 0

   #For each of the samples in the sample list
   for(d in sample_list_temp){
    
    #Add reads from samples_otu_complete [genus,sampleID]
    reads <- samples_otu_complete [c,d] + reads
    total <- samples_otu_complete["Total_Reads",d] + total
   }
   
   #Calc Rel abund (total number of reads / total number of reads)
   percent_obs <- (reads/total)
   
   #Create new header
   #Push value to new dataframe
   col_name <- paste(a,"-",b)
   samples_obs_abund_vari[c,col_name] <- percent_obs
  }
 }
}
#Output to text file
write.csv(samples_obs_abund_vari,"Data/ext_obs_abund_vari.csv")
```

#10. OTU Percent - Analysis Exp OTU Rel Abun % Diff by Control by Kit
```{r}
#Add expected frequencies and gram positive status to data table
for (a in control_types){
 
 #For each of the genus
 for (b in genus_expect){
  
  #Add gram pos/neg information
  samples_obs_abund_vari[b,"Gram"] <- control_expect[b,10]
  
  #Find value and pass to database
  value <- control_expect[b,a]
  
  if(is.na(value) | is.null(value)){
  } else{
   #Create new column
   col_name <- paste(a,"_Exp",sep="")
   samples_obs_abund_vari[b,col_name] <- value
  }
  }
}

#Output to text file
write.csv(samples_obs_abund_vari,"Data/ext_obs_abund_vari.csv")

#Start counter
count=1

#For each of the AC/ZE Controls
for (i in 8:17){
 
 #For each genus
 for (a in genus_expect){
  
  #Find the correct expected column
  col_name <- paste(control_types[count],"_Exp",sep="")
  
  #Find the value for the observe rel abund and the expected rel abund
  value1 <- samples_obs_abund_vari[a,i]
  value2 <- samples_obs_abund_vari[a,col_name]
  
  #Calculate percent difference
  if(is.na(value1) | is.null(value1)| is.na(value2) | is.null(value2)){
   value <- "NA"
  } else{
   value <- ((value1 - value2)/ value2)
  } 
  
  #Create new column name
  col_name <- paste((colnames(samples_obs_abund_vari[i])),"-Df",sep="")
  
  #Pass values
  samples_obs_abund_vari[a,col_name]<- value
  
 }
 
 #Move the counter to the next control type
 if(count==length(control_types)){
   count=1
 } else{count=count+1}
}

#Output to text file
write.csv(samples_obs_abund_vari,"Data/ext_obs_abund_vari.csv")
```

#11. OTU Percent - Analysis Unexp OTU Rel Abun by Control by Kit
```{r}
#Create full list of genus present
genus_present <- rownames(samples_otu_complete)
genus_expect<-as.character(genus_expect)

#Create list of unexpected
genus_unexpect <- genus_present[which(!genus_present %in% genus_expect)]

#Remove Total_reads from list
genus_unexpect <- genus_unexpect[genus_unexpect !="Total_Reads"]

#Create new dataframe
samples_obs_abund_vari_unexp <- data.frame()

#For each of the kits types
for (a in kit_types){
 
  #For each Genus in list
  for (b in genus_unexpect){
   
   #Reset Counter
   reads <- 0
   total <- 0
   
   #Create a list of samples that are of the specified control type
   sample_list_temp<- subset(samples_variables_f, Kit==a)
   sample_list_temp <- unique(sample_list_temp$NephID)
   
   #For each of the samples in the sample list
   for(c in sample_list_temp){
    
    #Add total reads from samples_otu_complete [genus,sampleID]
    reads <- samples_otu_complete [b,c] + reads
    total <- samples_otu_complete["Total_Reads",c] + total
   }
   
   #Calc Rel abund (total number of reads / total number of reads)
   #Push value to new dataframe
   percent_obs <- reads / total
   samples_obs_abund_vari_unexp[b,a] <- percent_obs
  }
}
#Output to text file
write.csv(samples_obs_abund_vari_unexp,"Data/ext_obs_abund_vari_unexp.csv")

#For each of the kits types
for (a in kit_types){
 
 #For each of the control types
 for(b in control_types){
  #Create a list of samples that are of the specified control type
  sample_list_temp<- subset(samples_variables_f, Kit==a & Type==b)
  sample_list_temp <- unique(sample_list_temp$NephID)
  
  #For each of the expected genus
  for(c in genus_unexpect){
   
   #Reset Counter
   reads <- 0
   total <- 0

   #For each of the samples in the sample list
   for(d in sample_list_temp){
    
    #Add reads from samples_otu_complete [genus,sampleID]
    reads <- samples_otu_complete [c,d] + reads
    total <- samples_otu_complete["Total_Reads",d] + total
   }
   
   #Calc Rel abund (total number of reads / total number of reads)
   percent_obs <- (reads/total)
   
   #Create new header
   #Push value to new dataframe
   col_name <- paste(a,"-",b)
   samples_obs_abund_vari_unexp[c,col_name] <- percent_obs
  }
 }
}
#Output to text file
write.csv(samples_obs_abund_vari_unexp,"Data/ext_obs_abund_vari_unexp.csv")
```

#12 Plot graph - Actual to Expected - Intrakit (Artificial Colonies)
For each kit, show variability, and map
```{r}
#Create list for plots, start counter
plot_list <- list()
i=1

for (a in control_types){
   #Create genus list of specfic control
   genus_temp <- control_expect[!is.na(control_expect[,a]),]
   genus_temp <- (genus_temp[,"Genus"])
   
   for (b in kit_types){
    temp<-data.frame()
    count=1
    
    ###Create database
    temp2 <- subset(samples_variables_f,Kit==b)
    temp2 <- subset(temp2,Type==a)
    
    #Subset, and create a sample list
    sample_list <- unique(temp2$NephID)
    sample_list
    
    #For each Genus in list, determine the samples value
    for (c in genus_temp){
     
     #For each of the samples in the sample list
     for(d in sample_list){
      temp[count,"Genus"]<-c
      temp[count,"Obs_Per"]<- samples_obs_abund[c,d]
      temp[count,"Sample"]<-d
      temp[count,"Exp_Per"]<-control_expect[c,a]
      count=count+1
     }
    }
    
    graph_title <- paste (a,": Intrakit Variablity -",b,sep = "")

    #Plot Data
    graph_new <- ggplot(data=temp,mapping=aes(x=Exp_Per,y=Obs_Per,color=Sample))+
     geom_point()+ #Can add points to graph
     geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
     labs(title=graph_title, x="Expected Percentage", y= "Actual Percentage")+
     geom_text(mapping=aes(x=Exp_Per,y=Obs_Per,label=Genus),check_overlap = TRUE)
    
    plot_list[[i]] = graph_new
    i=i+1
   }
}

#Start counters
count1=1
count2=1

#Save all plots
for (i in 1:(length(control_types)*5)){
 
 #ID which control is being run
 a <- control_types[count1]
 
 #Save the file
 file_name = paste("Data/Graphs/Intrakit Variability_",a,"-",kit_types[count2], ".tiff", sep="")
 tiff(file_name)
 print(plot_list[[i]])
 dev.off()
 
 #Move control counter, after 5 rounds (5 kits)
 #Reset the kit counter after 5 rounds
 if(i%%5==0){
  count1 = count1+1
  count2 = 1
  } else{
   count1=count1
   count2=count2+1
   }
} 

```

#13. Plot graph - Actual to Expected - All kits
```{r}
###############
#Zymo Ext
###############
#Create dataframe
zymo.kits <- samples_obs_abund_vari[,c(19,9,11,13,15,17)]
zymo.kits <- subset(zymo.kits,!(is.na(Zymo_Exp)))

#Check
rownames(zymo.kits)
colnames(zymo.kits)

#Create dataframe
temp<-data.frame()
count=1

#For each of the Kits
for (a in c(2:6)){
 count2=1
 
 #for each genus
 for(b in 1:nrow(zymo.kits)){
  temp[count,"Exp_Per"]<- zymo.kits[count2,"Zymo_Exp"]
  temp[count,"Genus"] <- rownames(zymo.kits[b,])
  temp[count,"Obs_Per"] <- zymo.kits[b,a]
  temp[count,"Kit"] <- colnames(zymo.kits[a])
  count=count+1
  count2=count2+1
 }
}
head(temp)

#Plot graph
graph_new<- ggplot(data=temp,mapping=aes(x=Exp_Per,y=Obs_Per,color=Kit))+
 #geom_point()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="Zymo.Extraction - Kit Comparison", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(x=Exp_Per,y=Obs_Per,label=Genus),check_overlap = TRUE)

file_name =("Data/Graphs/Kit Differences_ZE_.tiff")
tiff(file_name)
print(graph_new)
dev.off()

###############
#AC
###############
#Create dataframe
ac.kits <- samples_obs_abund_vari[,c(20,8,10,12,14,16)]
ac.kits <- subset(ac.kits,!(is.na(AC_Exp)))

#Check
rownames(ac.kits)
colnames(ac.kits)

#Create dataframe
temp<-data.frame()
count=1

#For each of the Kits
for (a in c(2:6)){
 count2=1
 
 #for each genus
 for(b in 1:nrow(zymo.kits)){
  temp[count,"Exp_Per"]<- ac.kits[count2,"AC_Exp"]
  temp[count,"Genus"] <- rownames(ac.kits[b,])
  temp[count,"Obs_Per"] <- ac.kits[b,a]
  temp[count,"Kit"] <- colnames(ac.kits[a])
  count=count+1
  count2=count2+1
 }
}
head(temp)

#Plot graph
graph_new<- ggplot(data=temp,mapping=aes(x=Exp_Per,y=Obs_Per,color=Kit))+
 #geom_point()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="Artificial Colony - Kit Comparison", x="Expected Percentage", y= "Actual Percentage")+
 geom_text(mapping=aes(x=Exp_Per,y=Obs_Per,label=Genus),check_overlap = TRUE)

file_name =("Data/Graphs/Kit Differences_AC_.tiff")
tiff(file_name)
print(graph_new)
dev.off()
```