---
title: "R Notebook"
output: word_document
editor_options: 
  chunk_output_type: console
---
#1. Load required libraries
```{r}
library("biom")
library(data.table)
library(ggplot2)
```

#2. Filtering of samples, control variable information
```{r}
#Read in database with variable information
samples_variables_complete <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Analysis\\ExpA_Run1&2\\Manifests\\sampledata_abrf.txt",sep="\t",header=TRUE)

#Read in database of failed samples
samples_failed <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Analysis\\ExpA_Run1&2\\TaxaCounts\\samplesfailed_q2.txt",sep="\t")

#Remove failed samples from info database
samples_filtered <-subset(samples_variables_complete, !(NephID %in% samples_failed$V1))

#Verify filtering - numbers should match
nrow(samples_variables_complete) - nrow(samples_failed)
nrow(samples_filtered)

#Only include Ext Controls
control_types <- c("RG","Study")
samples_variables_f <- samples_filtered[samples_filtered$Type %in% control_types,]

#Add sample ID's as row names
row.names(samples_variables_f) <- samples_variables_f$NephID
```

#3. OTU Counts by sample
Read in OTU information, and determine total counts for all samples
```{r}
#Read in OTU database at Genus level
samples_otu_complete <- read.csv("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Analysis\\ExpA_Run1&2\\TaxaCounts\\qiime2_taxa_counts_genus_silva_abrf.csv", row.names = 1)

#Create a list of all samples and all genus present in at least one sample
sample_list_complete <- colnames(samples_otu_complete)
genus_found <- rownames(samples_otu_complete)

#For each sample, determine the total read counts
for (a in sample_list_complete){
 #Set sum to 0
 sum=0
 
 #For each genus
 for (b in genus_found){
  #Add the number of counts together
  sum <- samples_otu_complete[b,a] + sum
 }
 
 #When complete, push the sum to a new "Total_Reads" row
 samples_otu_complete["Total_Reads",a] <- sum
}
write.csv(samples_otu_complete,"temp.csv")
```

#4. OTU Counts by Control Type
Determine genus count for each sample, average across control type
```{r}
#Create control sample list, control type list
sample_list_f <- unique(samples_variables_f$NephID)
control_types

#Genus count for each sample 
for (a in sample_list_f){
  count_present=0
  
  for (b in genus_found){
   value <- samples_otu_complete[b,a]
   
   if(value==0) {
    next
   } else{
     count_present = count_present +1
    }
  }
  samples_variables_f[a,"GenusObs"] <- count_present

 }
 
###Testing
write.csv(samples_variables_f,"temp1.csv")


#New Database to store final Genus information
samples_obs_count_type <- data.frame()
kit_types<-unique(samples_variables_f$Kit)

#Start counter
count_present=0

#For each control type
for (a in control_types){
 
 #For each kit
 for (b in kit_types){
  #Create temp database for kit and control type
  temp <- subset(samples_variables_f,Type==a)
  temp <- subset(temp,Kit==b)
  
  #Create sample list
  sample_list<- unique(temp$NephID)
  
  #For each of these samples
  for (c in sample_list){
   
   #Find observed value and add
   value <- temp[c,"GenusObs"]
   count_present=count_present+value
  }
  
  #Create new column and pass average
  col_name <- paste(a,"_",b,sep="")
  samples_obs_count_type["GenObs_Av",col_name]<-count_present/length(sample_list)
  
  #Reset counter
  count_present=0
 }
}
write.csv(samples_obs_count_type,"Data/study_avgenus.csv")
```

#5. OTU Percent - Analysis Genus OTU Abundance
```{r}
#Create new table
samples_obs_abund = data.frame()

##Run through each sample
for (a in sample_list_f){
 
 #Check for each of the expected genus
 for (b in genus_found){
  
  #Determine the reads for the genus, and the total reads for the sample
  reads <- samples_otu_complete[b,a]
  total <- samples_otu_complete["Total_Reads",a]

  #If value is NA (meaning that the genus was not found in sample), skip
  #If reads are >0 then add the read count to reads_total

  if(is.na(reads) | is.null(reads)) {
   next
  } else {
   percent_exp <- (reads/total)
   
   #Add percent to the samples_obs_abund database
   samples_obs_abund[b,a] <- percent_exp
  }
 }
}

#Output the table of all samples
write.csv(samples_obs_abund,"Data/study_obs_abund.csv")
```

#6. OTU Percent - Analysis Genus OTU Abundance by Control Type
```{r}
#New Database to store final Genus information
samples_obs_abund_vari <- data.frame()

#For each of the control types
for (a in control_types){
 
  #For each Genus in list
  for (b in genus_found){
   
   #Reset Counter
   reads <- 0
   total <- 0
   
   #Create a list of samples that are of the specified control type
   sample_list_temp<- subset(samples_variables_f, Type==a)
   sample_list_temp <- unique(sample_list_temp$NephID)
   
   #For each of the samples in the sample list
   for(c in sample_list_temp){
    
    #Add up the total percentages from samples_obs_abund [genus,sampleID]
    reads <- samples_otu_complete [b,c] + reads
    total <- samples_otu_complete["Total_Reads",c] + total
   }
   
   #average of genus level by taking total number of reads of that genus divded by total number of reads for samples included
   #Push value to new dataframe
   percent_obs <- reads / total
   samples_obs_abund_vari[b,a] <- percent_obs
  }
 }

#Output to text file
write.csv(samples_obs_abund_vari,"Data/study_obs_abund_vari.csv")
```

#7. OTU Percent - Analysis Exp OTU Rel Abun by Control by Kit
```{r}
#For each of the kits types
for (a in kit_types){
 
  #For each Genus in list
  for (b in genus_found){
   
   #Reset Counter
   reads <- 0
   total <- 0
   
   #Create a list of samples that are of the specified control type
   sample_list_temp<- subset(samples_variables_f, Kit==a)
   sample_list_temp <- unique(sample_list_temp$NephID)
   
   #For each of the samples in the sample list
   for(c in sample_list_temp){
    
    #Add total reads from samples_otu_complete [genus,sampleID]
    reads <- samples_otu_complete [b,c] + reads
    total <- samples_otu_complete["Total_Reads",c] + total
   }
   
   #Calc Rel abund (total number of reads / total number of reads)
   #Push value to new dataframe
   percent_obs <- reads / total
   samples_obs_abund_vari[b,a] <- percent_obs
  }
}
#Output to text file (adds new columns to prevoius version)
write.csv(samples_obs_abund_vari,"Data/study_obs_abund_vari.csv")

#For each of the kits types
for (a in kit_types){
 
 #For each of the control types
 for(b in control_types){
  #Create a list of samples that are of the specified control type
  sample_list_temp<- subset(samples_variables_f, Kit==a & Type==b)
  sample_list_temp <- unique(sample_list_temp$NephID)
  
  #For each of the found genus
  for(c in genus_found){
   
   #Reset Counter
   reads <- 0
   total <- 0

   #For each of the samples in the sample list
   for(d in sample_list_temp){
    
    #Add reads from samples_otu_complete [genus,sampleID]
    reads <- samples_otu_complete [c,d] + reads
    total <- samples_otu_complete["Total_Reads",d] + total
   }
   
   #Calc Rel abund (total number of reads / total number of reads)
   percent_obs <- (reads/total)
   
   #Create new header
   #Push value to new dataframe
   col_name <- paste(a,"-",b)
   samples_obs_abund_vari[c,col_name] <- percent_obs
  }
 }
}
#Output to text file (adds columns to file)
write.csv(samples_obs_abund_vari,"Data/study_obs_abund_vari.csv")
```

#8 - Not Used
```{r}
#Sort genus list alphabetically
genus_found<-sort(genus_found)

for (a in control_types){
   for (b in kit_types){
    temp<-data.frame()
    count=1
    
    ###Create database
    temp2 <- subset(samples_variables_f,Kit==b)
    temp2 <- subset(temp2,Type==a)
    
    #Subset, and create a sample list
    sample_list <- unique(temp2$NephID)
    sample_list
    
    #For each sample
    for (c in sample_list){
     
     #Grab value at each genus
     for (d in genus_found){
      value <- samples_obs_abund[d,c]
      
      if(is.na(value) | is.null(value) | value<0.0000000001){
       next}
      else{
       temp[d,"Abundance"]<-value
       temp[d,"Genus"]<-d
      }
    
     }
     write.csv(temp,"temp4.csv")
    temp<-data.frame()
   }
   }
}
temp<-read.csv("temp4.csv",header = TRUE)
temp

library(reshape)
datm <- melt(cbind(dat, ind = rownames(dat)), id.vars = c('ind'))

library(scales)
ggplot(datm,aes(x = variable, y = value,fill = ind)) + 
    geom_bar(position = "fill",stat = "identity") +
    # or:
    # geom_bar(position = position_fill(), stat = "identity") 
    scale_y_continuous(labels = percent_format())

```